### 2012-02-27 eswc-2012/workshop.tex
one question; more links
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@119 43ef1190-90fd-4dd7-a769-d0121f02b80e

\documentclass{llncs}
%\documentclass{article}
\usepackage{cite}

\title{Practical implications of HATEOAS on a bridge between LOD and SPARQL}
\titlerunning{HATEOAS and a bridge between LOD and SPARQL}
\author{Kjetil Kjernsmo\inst{1}}
\institute{Department of Informatics,
Postboks 1080 Blindern,
0316 Oslo, Norway
\email{kjekje@ifi.uio.no}}

\begin{document}

\maketitle

\begin{abstract}
This paper makes two points: 1) SPARQL is a necessity for non-trivial
navigation in Linked Data, 2) The community needs to pay attention to
the HATEOAS constraint of REST and its practical implications for
making a bridge between LOD and SPARQL possible.

\end{abstract}

\section*{BRAINDUMP STATUS}

\section{Introduction}

Some bloggers in the developer community, see e.g.~\cite{sunsetonmvc},
is currently predicting a paradigm shift in common practices, prompted
by the trend towards richer clients. This has caused a departure from
the time-tested Model-View-Controller pattern that has been the
dominant paradigm in web development for more than a decade.

Semantic Web services is likely to be an important source of data for
future applications, but for foreseeable future just one of many. A
key promise of Linked Open Data (LOD) in particular, and the Semantic
Web in general is the ability to integrate many data sources
easily. However, this integration requires links and it requires that
the resulting graph can be traversed by reasonable means. While links
are quite abundant across the LOD cloud, this paper will argue that
key links are missing to make it possible to traverse the resulting
graph. Moreover, we will argue that this shortcoming is due to that
the community does not adequately take into account the constraint
known as ``Hypermedia as the Engine of Application State'' (commonly
abbreviated HATEOAS) from the REST architectural style, see
\cite{fielding}~Chapter~5.

The HATEOAS constraint requires that the application can navigate from
one resources to another using the hypermedia links in the present
resource \emph{only}, it should not require any out-of-band
information to do so. Thus, a RESTful API should not require prior
knowledge of a URI structure, but one may define the API in terms of
URI structure as long as the same information can be gleened from the
resources. Nor does it require that all resources can be reached from
any other resource. As such, the constraint is not very strict, one
could argue that as long as there are some links, the API satisfies
the HATEOAS constraint.

Since RDF is built on URIs, many of which can be dereferenced to
obtain further links, it lends itself well to RESTful APIs, but it is
debateable whether an API can be said to be RESTful if the links
doesn't enable any useful interactions. Therefore, whether the HATEOAS
constraint is satisfied must be judged on the basis of the practical
applications the API enables.

The LOD Cloud consists of well-linked datasets and to find simple
answers by simply dereferencing URIs, parse the returned content and
address it by conventional methods provided by the programming
language of choice, and this pattern has proved very useful. However,
for slightly more complex applications, one has to traverse a
relatively large graph to find the data one requires. In most cases,
one would need to match the graph with a pattern, which is essentially
what the SPARQL query language does. The main problem is that there
may not be a link present that makes it possible for the application
to find the query endpoint without requiring out-of-band information,
which implies that non-REST interactions are required. To remedy this
situation, we would need to add more links.

\section{Questions for discussion}

\begin{enumerate}
\item TimBL suggests in the Linked Data Design
  Issue\footnote{\url{http://www.w3.org/DesignIssues/LinkedData}} the
  use of ``Minimum Spanning Graph'', what other triples would humans
  and machines find useful?

\end{enumerate}

%\bibliographystyle{plain-csmin}
%\bibliographystyle{jbact}
%\bibliographystyle{splncs03}
%\bibliography{selectivity,federation,benchmarks,optimization}

\end{document}