### 2012-10-19 dissertation/introsota.tex
Add stub on dissertation
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@392 43ef1190-90fd-4dd7-a769-d0121f02b80e

\section{State of the art}

The state of the art.