### 2012-10-19 dissertation/introsota.tex
Fix citations; new label on PRM selectivity
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@393 43ef1190-90fd-4dd7-a769-d0121f02b80e

\section{State of the art}

The idea to use Statistical Relation Learning techniques for
selectivity estimation first originated in \cite{selectivityPRM}.