### 2012-10-23 dissertation/introproblem.tex
Add problem
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@395 43ef1190-90fd-4dd7-a769-d0121f02b80e

\section{Problem description}

The main topic of the present dissertation is SPARQL
Federation\todo{further discuss different interesting directions with
  SPARQL federation}.