### 2012-10-25 dissertation/tmpsolutions.tex
Discussion of solutions
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@407 43ef1190-90fd-4dd7-a769-d0121f02b80e

\section{Possible directions of work}

\emph{This section is a temporary discussion area}

We need to discuss the future strategy for this work. I think the main
discussion point is the tension between elaborately develop theory
that then can be applied to derive publishable results, vs. the
practical time constraints of Ph.D. work. Another issue is that I'd
like to be mindful of the realities of adoption in standards and
adoption of standards.