### 2013-06-20 exp1/rebuttal.txt
rebuttal
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@592 43ef1190-90fd-4dd7-a769-d0121f02b80e

We would like to thank the anonymous reviewers for their thoughtful
reviews. Your points are well taken and we shall work to incorporate
your advices to the final version.

First, benchmarking and designed experiments serve to answer different
questions. Since a designed experiment requires parameterization of
language and data "features", natural benchmarks such as the DBPedia
Benchmark, cited in the paper, will remain a better judge of the
performance of typical queries in some typical situations. However, it
cannot, as we noted in the three points on the first page, answer
scientifically relevant questions on e.g. falsifiability of more general
assertions about performance.

One reviewer notes that it will be tricky to extend the experiment to
real federated SPARQL querying. Indeed, we believe this to be the work
of more than one Ph.D. Thus, this paper is to encourage a collective
effort of the community. It is not unlikely that to attain this goal,
progress beyond the state-of-the-art in statistics is also required.

Two reviewers point out that the introduction to DoE should be
improved. This point is very valid. We envision this paper to be a
"first contact" with DoE for many readers, so the presentation is
especially important, and your suggestions for improvements are very
welcome.

Obviously, we also acknowledge that the "toy example" nature of the
paper is a clear drawback. We are quite confident that this
methodology will be a clear improvement over "constructed" benchmarks,
like for example the now classical Berlin SPARQL Benchmark with
relatively modest means, but we acknowledge that we have not attempted
to prove this notion. We shall attempt to provide at least anecdotal
evidence based on the current example for the camera ready
version. Yet, we are of the opinion that if we had attempted a full
replacement of the larger benchmarks in this paper, the discussion
would have been significantly harder to follow.

As an expansion of the introduction DoE requires more space, one
reviewer's suggestion to drop Table 3 is welcome. However, Table 3 was
not intended as a preliminary investigation, but rather an
illustration of the concept of "lurking variables", which is
conceptually important. We would be interested in hearing the
reviewer's opinion on the importance of this point.

In conclusion, there are no major confusion points to be cleared, the
reviews are helpful and if the paper is accepted, we shall seek to
incorporate their suggestions to the extent possible within the page
limit.