### 2015-02-19 sanity/rebuttal.txt
mention federated SPARQL
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@1032 43ef1190-90fd-4dd7-a769-d0121f02b80e

I would like to thank the reviewers for their extensive reviews.

I would first like to address the main point of critisism voiced
mainly by Review 4 but also by Review 2, that the survey is of a very
limited "Semantic Web" and so is unlikely to be generalizeable.

This critisism is very valid and taken to heart. To understand the
constraints of the study, it is first important to note that it is
directed at the community that convenes at ESWC. Also, such a paper
could address what this community does, what it can do, and what it
should do. I have consciously avoided strong "should do"
recommendations, except in the case where current practice is clearly
misguided. Most recommendations are therefore of the much softer "can
do" type, but the overall focus of the paper is on what the community
does.

Consequently, this survey is not meant to be generalizeable beyond the
current state of the art. I find it quite likely that the current
state must change in order for the Semantic Web to be successful, and
if the Semantic Web succeeds, the current conclusions will most
certainly not hold. In particular, there is nothing on the Web today
that supports query evaluation in the way that a large-scale
deployment of federated SPARQL would. The scope of the survey is a
snapshot, confined to what this community can reasonably do something
about, and currently, we are such a tiny fraction of the Web, that we
have to understand the Web better to succeed.

This survey is not intended as means to understand what the rest of
the Web is doing right and wrong, that wisdom is available elsewhere
in the broader literature, but since I believe a survey paper should
focus mostly on "does", it was not within my scope, nor the pagelimit,
to cite. This wisdom was therefore only cited from the standards.

The intention is mostly to have empirical research to see if 
statements like "it is no point in using caches when evaluating SPARQL
queries, since no endpoints are exposing them" have merit.

Review 1 asked us to discuss the importance of caching metadata for
query processing. I believe this is inappropriate for two reasons, one
is the "does" focus of the paper, but more importantly, this is
largely unknown at present, and it cannot be inferred from a survey
alone. The study by Lampo et al, cited among related work, provides a
sobering account, and cautions us to not oversell the importance of
caching for query processing.

As this is an broad area of critisism from several reviews, I
shall attempt to condense the above discussion into a concise
clarification in the paper.

Review 4 pointed out a flaw in that the Surrogates header was wrong. I
acknowledge the error, and reran the spider recording the
Surrogate-Capability header instead. It was not found, and therefore
the flaw does not influence the conclusions, but I'll fix this.

I note that the paper is very close to the page limit, so I would have
to remove parts of the paper to be able to expand on other parts. In
fact, I removed discussions relevant to the questions raised by Review
3 from the initial draft as they felt to deviate from the focus too
much.

Another example of things I removed was a list of the contributions,
as requested by Review 6, as I felt it was merely a reiteration of the
objectives that didn't not serve to clarify the paper, for any other
reason than convention.

Review 6 suggests that I use a logarithmic scale in some of the
figures. I initially did this, but felt that they didn't effectively
convey the essence, and that the problem is that our time and calendar
doesn't follow linear progression either. Therefore, one is left with
the choice of a classification that doesn't make intuitively sense or
one that doesn't make sense mathematically. I'll try out the
suggestion of combining the two.

I'll discuss these issues with local faculty to find a good solution.

Review 6 also suggests that I restructure the paper, and focus the
discussion. With a focused discussion, it may be room to do some
expansions of other points. I'll try my best, but it does not seem
straightforward to include all the elaborations requested by the
reviewers within the page limit.

I'm not sure how to address the comment of Review 1 that the survey is
weak because it is based on only two documents. I assume they refer to
the two RFCs. First, it is not entirely true, it also references other
documents, only to show it is not widely used. Moreover, the two
documents defines the entire standards-based caching infrastructure of
the Web, as such, the survey is exhaustive.

Review 6 requests data to be published. I shall publish the data and
the code in a tar archive.

The reviewers provided detailed editorial comments that I shall take
into account when creating a camera ready version. I'll try to find a
native English writer to review the paper.