### 2015-02-21 sanity/rebuttal.txt
Shorten
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@1034 43ef1190-90fd-4dd7-a769-d0121f02b80e

I would like to thank the reviewers for their extensive reviews.

I would first like to address the main point of critisism voiced
mainly by Review 4 but also by Review 2, that the survey is of a very
limited "Semantic Web" and so is unlikely to be generalizeable.

This critisism is very valid and taken to heart, and I should strive
to make it clearer why this is the scope I chose to work within.
Indeed, this Semantic Web is a mere fraction of the open data web,
where e.g. JSON APIs are common. Coffee break discussions around
caching have often stranded on the lack of data around what the
community as a whole actually does, and therefore, that is the main
focus of this paper. What it can do is given some space, but a
discussion of what it should do is consciously avoided, except in the
case where current practice is clearly misguided. Since the paper
already fills the pagelimit, I have only cited the foundational
documents for wisdom on what is thought to be important for the Web, a
further discussion would be beyond the available space, and probably
also be better suited for a different conference.

Review 1 asked us to discuss the importance of caching metadata for
query processing. This is the focus our research, but this is
largely unknown at present, and it cannot be inferred from a survey
alone. The study by Lampo et al, cited among related work, provides a
sobering account, and cautions us to not oversell the importance of
caching for query processing.

Review 4 pointed out a flaw in that the Surrogates header was wrong. I
acknowledge the error, and will rerun the spider.

I note that the paper is very close to the page limit, and amongst the
things that were removed from the initial draft are discussions of the
questions raised by Review 3 and a list of the contributions, as
requested by Review 6. Review 6 suggest that I focus the discussion,
which could reduce the length somewhat, but I would welcome
suggestions on what else could be trimmed, to make room for
elaborations suggested in the reviews.

I initially used logarithmic scale as suggested by Review 6, for the
reasons they mention. I found the intuitive interpretation of the
figures to be very difficult, since the time and calendar scale is as
it is. The suggestion to combine the two is interesting, and I will
see if provides any benefit if the paper is accepted.

I'm not sure how to address the comment of Review 1 that the survey is
weak because it is based on only two documents. I assume they refer to
the two RFCs. First, it is not entirely true, it also references other
documents, only to show that they are not widely used. Moreover, the
two documents defines the entire standards-based caching
infrastructure of the Web, as such, the survey is exhaustive.

Review 6 requests data to be published. I shall publish the data and
the code in a tar archive.

The reviewers provided detailed editorial comments that I shall take
into account when creating a camera ready version. I'll try to find a
native English writer to review the paper.