### 2015-02-21 sanity/rebuttal.txt
shorten
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@1036 43ef1190-90fd-4dd7-a769-d0121f02b80e

I would like to thank the reviewers for their extensive reviews. The
many editorial advices will certainly be taken into account without
further comment.

The critisism that I only surveyed a small part of the Web and so, the
conclusions are not generalizeable is valid, and I should strive to
make it clearer why this is the scope I chose to work within. Indeed,
this Semantic Web is a mere fraction of the open data web, where
e.g. JSON APIs are common. ISWC/ESWC coffee break discussions around
caching have often stranded on the lack of data around what the
community as a whole actually does, and therefore, that is the main
focus of this paper. The recommendations are meant as suggestions on
what the community may do, but a discussion of what it should do is
consciously avoided, except in the case where current practice is
clearly misguided. Since the paper already fills the pagelimit, I have
only cited the foundational documents for wisdom on what is thought to
be important for the Web, a further discussion would be beyond the
available space.

Review 1 asked us to discuss the importance of caching metadata for
query processing. This is the focus our research, but this is
largely unknown at present, and it cannot be inferred from a survey
alone. The study by Lampo et al, cited among related work, provides a
sobering account, and cautions us to not oversell the importance of
caching for query processing.

Review 4 pointed out a flaw in that the Surrogates header was wrong. I
acknowledge the error, and have rerun the spider.

I note that the paper is very close to the page limit, and amongst the
things that were removed from the initial draft are discussions of the
questions raised by Review 3 and a list of the contributions, as
requested by Review 6. Review 6 suggest that I focus the discussion,
which could reduce the length somewhat, but I would welcome
suggestions on what else could be trimmed, to make room for
elaborations suggested in the reviews.

I initially used logarithmic scale as suggested by Review 6, for the
reasons they mention. Then, I found the intuitive interpretation of
the figures to be very difficult, since the time and calendar scale is
as it is. The suggestion to combine the two is interesting, and I will
see if it provides any benefit if the paper is accepted.

I'm not sure how to address the comment of Review 1 that the survey is
weak because it is based on only two documents. I assume they refer to
the two RFCs. First, it is not entirely true, see Table 1. Moreover,
the two documents define the entire standards-based caching
infrastructure of the Web, as such, the survey is exhaustive.