### 2015-03-11 sanity/paper1.tex
Long comment on coverage and reduction
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@1061 43ef1190-90fd-4dd7-a769-d0121f02b80e

\documentclass{llncs}
%\documentclass{sig-alternate}
\usepackage[english,british]{babel}
%\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{cite}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{hyperref}

\newcommand{\rdfterm}[1]{\texttt{#1}}
\newcommand{\pvalue}{\textit{p}-value\ }
\newcommand{\httph}[1]{\texttt{#1}}
\newcommand{\todo}[1]{\ensuremath{^{\textrm{\tiny{TODO}\normalsize}}}\footnote{\textbf{TODO:}~#1}}

\title{A survey of HTTP caching implementations on the open Semantic Web}
\author{Kjetil Kjernsmo}
\institute{Department of Informatics,
Postboks 1080 Blindern,
N-0316 Oslo, Norway \email{kjetil@kjernsmo.net}}

%\subtitle{---Unpublished Working Draft, do not circulate---}

\begin{document}
\maketitle

\begin{abstract}
Scalability of the data access architecture in the Semantic Web is
dependent on the establishment of caching mechanisms to take the load
off of servers.  Unfortunately, there is a chicken and egg problem
here: Research, implementation, and evaluation of caching
infrastructure is uninteresting as long as data providers do not
publish relevant metadata.  And publishing metadata is useless as long
as there is no infrastructure that uses it.

We show by means of a survey of live RDF data sources that caching
metadata is prevalent enough already to be used in some cases.  On the
other hand, they are not commonly used even on relatively static data,
and when they are given, they are very conservatively set. We point
out future directions and give recommendations for the enhanced use of
caching in the Semantic Web.
\end{abstract}

\section{Introduction}

Caching has been given a prominent place in the foundational documents
of the World Wide Web. Out of the 6 documents that make up the HTTP
1.1 standard, RFC7234 \cite{rfc7234} is entirely devoted to the topic,
and RFC7232 \cite{rfc7232} defines conditional requests, which is also
important when constructing caches. As the former notes:
\begin{quote} 
  The goal of caching in HTTP/1.1 is to significantly improve
  performance by reusing a prior response message to satisfy a current
  request.
\end{quote}
Furthermore, the Architecture of the World Wide Web
\cite{Jacobs:04:AWW} discusses caching throughout, and the definition
of the Representational State Transfer (REST) architectural style
\cite{Fielding_2000_Architectural-Styles} is partly motivated from the
requirement to implement efficient caching. We also note that caching
in the Internet infrastructure, through so-called Content Delivery
Networks, is both a large business area and something that could
provide great value to the Semantic Web.

If used correctly, caching mechanisms will reduce the need to make
HTTP requests, reduce lookups to the backend systems, reduce the need
to make repetitive computations, enable sharing of responses in
Internet infrastructure, improve uptime and reduce latency since
requests may be answered closer to the client.

In spite of this, we have not seen it in widespread use in the
Semantic Web, and therefore we decided to conduct a survey to
investigate the actual compliance to RFC7234 and RFC7232. 
The objectives of this paper are:
\begin{enumerate}
\item Understand the actual usage rather than rely on anecdotal
  conceptions.
\item Encourage the implementation of these mechanisms in Semantic Web
  infrastructure.
\item Point out future research directions.
\end{enumerate}
The contributions of this paper are to meet these objective by means
of a survey that shows that while the uptake has been moderate,
practical benefits may be realized already. Based on this survey as
well as practical experience, we point out future research directions
as well as some recommendations for deployed implementations.

We note that caching is not only useful for long-living information
resources, even though that may be the most important use. If a resource
is frequently requested, it may make sense to cache it even though it
may be fresh for only a very short period.

Caching may be deployed at several different levels: An HTTP cache may
be in a reverse proxy close to the server, in which case it may have
much in common with a conventional database cache. It may also be
anywhere between a server and a client, in which case it may be
shared, i.e. it may cache responses from several servers served to
several clients. Another example is an institutional forward proxy,
which are close to several users. Finally, the User Agent may
implement a private cache for its user at the client side.

\subsection{HTTP Caching standards}

As mentioned, the two documents from the HTTP 1.1 standards suite that
are relevant for this study are RFC7234, which is named ``Caching''
and RFC7232, named ``Conditional Requests''. The main difference is
that the caching standard defines when a response may be reused
without contacting the origin server at all, whereas the conditional
requests define how to validate a response by contacting the origin
server. The two can be combined: Clients and proxies may use the
latter to revalidate a response that has been cached based on the
former.

RFC7234 defines two important headers: \httph{Expires}, whose value is
a date and time of when the response is considered stale, and
therefore should not be used; and \httph{Cache-Control}, which allows
detailed control of the cache, including a \httph{max-age} field,
which gives the time in seconds for how long the the response may be
used from the time of the request. \httph{max-age} takes precedence
over \httph{Expires}. In the following \emph{freshness lifetime} is
understood as the number of seconds that the response may be used
without contacting the origin server. Ideally, the calculation of the
freshness lifetime should be based of the above, we therefore shall
refer to this as ``standards-compliant caching''. However it can also
be based on heuristics, and Section~4.2.2 in RFC7234 provides some loose
constraints for such practice, as well as a suggestion for a useful
heuristic. This heuristic is based on a fraction of the time lapsed
between the current time and the modification time given in the
\httph{Last-Modified} header. This approach therefore still requires
the Web server to be cooperative to be successful. Commonly, Web
servers can track this, for example if RDF is served from a file
system the file modification time is used.

RFC7232, on the other hand, defines a protocol for asking the server
if the cached response is still fresh using conditional requests. This
doesn't burden the content provider with the task of estimating the
time to live beforehand, but then, it must be able to answer if the
resource has changed less expensively than serving the entire
response. Either of two headers must be set by the server to achieve
this: \httph{ETag}, which sets an opaque identifier for the response,
or \httph{Last-Modified} which gives the time and date of the last
modification of the resource. Clients may use these values to retrieve
the updated response only if the resource has changed. Other relevant
headers are described in a companion technical report
\cite{kjernsmo_add_survey_2015}.
% Table~\ref{tab:headers} in Appendix~\ref{app:fetcher}.
\todo{More on HTTP}

RFC7234 provides detailed control of caching, and caching may also be
prohibited by the server, either by setting a non-positive freshness
lifetime or explicitly using a \httph{no-store} control field.

In this paper, we study to what extent SPARQL endpoints, vocabulary
and data publishers support these standards.\todo{link to dump}

\section{Related work}

We are not aware of any surveys of this type. Although the database
literature is rich with query cache literature, it is mostly relevant
to what would happen within the server or between the server and a
reverse proxy, which is opaque to the Internet, and therefore not of
our concern. For the same reason, caching that happens within the
SPARQL engine is not our concern.

The Dynamic Linked Data Observatory (DyLDO) \cite{dyldo2} performed,
and continues to do so, monitoring of parts of the Linked Open Data
Cloud to determine dynamicity characteristics of Linked Data. Caching
is one of their motivations, but they have not published statistics on
HTTP headers.

Linked Data Fragments is claimed in \cite{ldf1} to take advantage of
caching and contrasts this with the unavailability of SPARQL query
caches, and it is claimed that this is an architectural problem. In
\cite{hogan2014paths}, the authors examine cacheable as one of the
desiderata for sustainable data access. They claim, without further
justification, that SPARQL isn't cacheable.

In \cite{sparqlproxy} the authors implemented a reverse proxy that
controlled the changes to the dataset, and therefore could make sure
the proxy had all the information needed to determine freshness. We
are interested in the situation where the changes cannot be
controlled.

In \cite{umbrich2012hybrid}, the term caching was used in a different
sense than we do, they rather prefetched an entire dataset to a local
store and based on heuristics tried to determine which parts of the
query should be evaluated remotely and locally. \cite{lampo2011cache}
explored when caching had a positive effect on complex SPARQL
queries.

In the broader Web literature, \cite{ager2010revisiting} analysed the
value of caching based anonymized traces of actual Web usage at a
major Internet Service Provider. They find that while caching often
yields little benefit when content is user-generated, they still find
some potential.

While these studies have little overlap with the present paper, they
underline the importance of understanding the current deployment and
future potential. In some of the related work, it is shown that
caching does not necessarily give tangible benefits. Yet, we shall
assume that sharing the metadata required for caching outside of the
server is desirable, and that it is possible in most cases. We shall
see that it most likely will be beneficial in cases that do not
benefit from caching today.

\section{Methodology}
We want to find information resources on the Web, and examine HTTP
headers that may allow caching. To do this, we perform GET requests on
SPARQL endpoints, vocabularies, dataset descriptions and other
resources and record headers recommended by current standards, as well
as obsoleted and non-standard headers. Additionally, we examine the
triples in the returned information resources to see if there is
information there that may be used to calculate heuristic freshness.

We make several approaches to ensure that we visit a large and
representative section of the open Semantic Web. We take SPARQL
Endpoints from the SPARQLES survey \cite{buil2013sparql}, vocabularies
from Linked Open Vocabularies (LOV) \cite{lov2} and prefix.cc, and
we augment these data with spidered data from the Billion Triple
Challenge (BTC) 2014 \cite{btc-2014} dataset.

We used SPARQLES survey list of SPARQL endpoints as of 2014-11-17, and
filtered out those deemed unresponsive. This resulted in a
list of 312 endpoints.

To examine as many different implementations and hosts as possible, we
noted that the Billion Triple Challenge 2014 \cite{btc-2014} dataset
consisted of a 4~GTriple corpus of spidered Web data. This was seeded
from datahub.io (aka CKAN), as well as other sources. To compile a list of
candidates to further examine, we performed a series of data reduction
steps, manually inspecting the result between each step. The details
of this process is given in the companion technical report
\cite{kjernsmo_add_survey_2015}.% Appendix~\ref{app:reduction}.

The end result of this process is a list of 3117 unique hosts, for
each several resources would be visited, some several times, as they
may host SPARQL endpoints, vocabularies, or other information
resources, by a spider detailed in the companion technical report
\cite{kjernsmo_add_survey_2015}, resulting in 7745 requests, done on
2015-01-02.

This results in an NQuads file per host, which is then loaded into a
Virtuoso-based SPARQL endpoint for analysis by using the statistics
system R\cite{kn:r} in the following section.\todo{Reorg, rev 6}

A key challenge to the validity of the survey is biases that may be
introduced by the coverage and then the data reduction. The breadth of
the Semantic Web is derived mainly from the BTC2014 crawl. While
LODstats\footnote{\url{http://stats.lod2.eu/}} has presently seen an
order of magnitude more triples, the number of error-free datasets
were at the time of this writing 4442. We work under the assumption
that cache headers are set mostly on a per-host basis, and if this
assumption is valid, sampling a URL per host is sufficient. LODstats
do not report per-host statistics, but often one host will host
several datasets. Another recent crawl was reported by
\cite{stateoflod2014}. It is not clear how many hosts were crawled, but the
number of triples in the crawl is much smaller than that of
BTC2014. It is therefore a fair assumption that BTC2014 fairly well
represent the breadth of the Semantic Web, momentarily at least.

As for the coverage of vocabularies, we realize that LODstats might
have been a better source, but we have been able to select the most
important, and not far inferior in number. The number of SPARQL
endpoints found in SPARQLES is larger than LODstats, and we also
looked for further endpoints both in the BTC2014 and our own crawl,
finding only a few. If endpoints went underdiscovered, then there is a
discovery problem that is beyond this survey to rectify.

The data reduction that was subsequently done was mainly done to
eliminate errors. We have not investigated biases that may be
introduced by discarding momentarily dysfunctional parts of the
Semantic Web, but we investigated whether the freshness lifetimes
reported in the case of certain errors were distributed differently
from those that returned a valid response, the companion technical
report \cite{kjernsmo_add_survey_2015}. We found that they were, but
we have assumed that this is due to that errors are configured to be
cached differently, which we know from experience is common
practice. The following analysis is based on valid responses.

\section{Analysis}

\subsection{Different server implementations}

First, we investigated whether certain server implementations provided
better support for caching than others. To do this, we formulated
SPARQL queries to examine the \httph{Server} headers of successful
responses, with optional clauses matching the standards-compliant
computed freshness lifetime (which is the ideal) as well as whether
the response had other indications of caching-related metadata that
may assist caching such as modification time, certain predicates, etc.

For each unique \httph{Server} header, we found the ones where
\emph{all} responses had a freshness lifetime or other usable
metadata. For the former, this amounted to 22 servers, which are
listed in Table~\ref{tab:servers}.  70 servers always responded with
usable metadata. Inspecting the values we find the well-known Virtuoso
and Callimachus servers, as well as the Perl module
RDF::LinkedData\footnote{See
  \url{https://metacpan.org/pod/RDF::LinkedData}. It is also present
  in Debian/Ubuntu as \texttt{librdf-linkeddata-perl}}, which is
developed by and runs on a server operated by this author. Apart from
those, we see that the they reveal very little about the RDF-specific
parts of the underlying server implementation. A quick inspection of
all \httph{Server} headers confirmed that few reveal any further
detail.

\begin{table}[h]
  \caption{\httph{Server} headers for hosts that enabled a freshness
    lifetime to be computed for \emph{all} requests.}\label{tab:servers}
\begin{center}
\begin{tabular}{rl}
  \hline
  1 & DFE/largefile \\ 
  2 & git\_frontend \\ 
  3 & nginx/1.3.9 \\ 
  4 & thin 1.6.0 codename Greek Yogurt \\ 
  5 & Oracle-Application-Server-10g/10.1.3.4.0 Oracle-HTTP-Server [\ldots]\\
  6 & Oracle-Application-Server-10g/10.1.3.4.0 Oracle-HTTP-Server [\ldots]\\
  7 & TwistedWeb/8.2.0 \\ 
  8 & RDF::Endpoint/0.07 \\ 
  9 & Jetty(6.1.26) \\ 
  10 & nginx/1.6.1 \\ 
  11 & Jigsaw/2.3.0-beta3 \\ 
  12 & Apache/2.2.9 (Win32) PHP/5.2.6 \\ 
  13 & Apache/2.4.10 (Unix) mod\_fcgid/2.3.9 \\ 
  14 &  \\ 
  15 & GFE/2.0 \\ 
  16 & RDF::LinkedData/0.70 \\ 
  17 & Apache/2.2.17 (Unix) mod\_wsgi/3.3 Python/2.6.6 \\ 
  18 & Virtuoso/07.10.3211 (Linux) i686-generic-linux-glibc212-64  VDB \\ 
  19 & Apache/2.2.24 (Unix) mod\_ssl/2.2.24 OpenSSL/0.9.8y \\ 
  20 & Apache/2.2.22 (Fedora) \\ 
  21 & INSEE \\ 
  22 & GitHub.com \\ 
   \hline
\end{tabular}
\end{center}
\end{table}

For a more systematic approach, we wish to test the hypothesis that
some servers are better configured to support caching than others.
Using the methodology given in the companion technical report
\cite{kjernsmo_add_survey_2015}, we find in
both the cases of standards-compliant freshness lifetime and for the other
usable metadata, the test reports \pvalue = 0.0001. We can conclude
that it is highly likely that some servers are better at exposing
cache headers than others. Unfortunately, since most \httph{Server}
headers only contain generic values, little can be learnt about these
implementations. We note, however, that DBPedia exposes
standards-compliant freshness lifetime of 604800 seconds (i.e. 1 week) for
both LOD and SPARQL endpoints.\todo{elaborate}

\subsection{Other caching headers}

The companion technical report \cite{kjernsmo_add_survey_2015}, lists other caching
headers than those mentioned in the introduction. We found
\httph{Pragma} (archaic HTTP 1.0 header) in 287 responses, but except
for two hosts, where they were superfluous, they were only used to
prohibit caching. \httph{Surrogate-Capability} were not observed.

\subsection{Distribution of freshness lifetime}

We obtained a successful response from a total 2965 information
resources, either with SPARQL results or RDF data. A successful
response is rather strictly defined, not only must there be a
successful HTTP response after redirects are resolved, the response
must also return a valid RDF media type (unless it is a SPARQL result)
and the response must parse into an RDF model. We have given priority
to survey many hosts since configuration usually doesn't differ much
across a host, especially since it also captures different types of
resources. It is therefore acceptable that the number of resources is
relatively small.

Since we are interested in the properties of valid response, including
examining some of the RDF contained in them, and that web servers may
be configured to instruct clients and proxies to cache errors
differently, we will study the statistical properties of valid
responses.

\subsubsection{Standards-compliant caching headers}

Of the 2965 resources, 405 returned valid headers, but 114 did so to
prohibit caching of the response, and 3 contained conflicting
headers, i.e. set a freshness lifetime, but also prohibited
caching. In most cases, \httph{Cache-Control} and \httph{Expires} both
occurred, but the former is more common than the latter in the cases
where only one of them occur.  Additionally, 269 resources had a
\httph{Cache-Control} header to control other aspects of caching than
lifetime, i.e. to say that only private caches may use the response,
that the cache must be revalidated, or to prohibit caching. 
Note that the freshness lifetime is 0 whenever caching is prohibited.

In Figure~\ref{fig:hardall}, there is a barplot where
the freshness lifetime is grouped in bins. We see that in these
categories, the most common is to prohibit caching. Nevertheless, many
also declare a standards compliant freshness lifetime in minutes to
days\todo{explain number of occur; fix labels}\todo{qqplot in report}.

\begin{figure}[ht]
  \centerline{%
    \includegraphics[width=.9\textwidth]{hardall.pdf}}
  \caption{Barplot counting all standards-compliant freshness lifetimes
    found, with the percentage of occurrences indicated on the
    bars. The first bin are the cases where caching is explicitly
    prohibited. The next bins are for lifetimes, where the values are
    grouped if they are on the order of seconds, minutes, hours, etc,
    i.e. the second bin counts the lifetimes in the interval [1,59]
    seconds, etc.}
  \label{fig:hardall}
\end{figure}

In Figure~\ref{fig:hardtable}, we have broken this
up by the type of resource that was accessed, i.e. SPARQL endpoints,
vocabularies, dataset descriptions or unclassified information
resources. First, we note that it seems like the distribution of
freshness lifetime is quite different for the different types, an
observation that is also supported by a similar hypothesis test as
above, with a \pvalue = 0.00001. We also note that it is often
prohibited to cache dataset descriptions. This is odd, since
statistics about datasets is usually costly to compute and should be
cached. The VoID specification \cite{voidnote} also notes that the
statistics are considered estimates.

\begin{figure}[ht]
  \centerline{%
    \includegraphics[width=.9\textwidth]{hardtable.pdf}}
  \caption{Mosaic Plot. On the vertical axis, the size of the boxes
    are determined by the fraction of the types of resources. On the
    horizontal axis the width of the boxes is proportional to the
    total counts, using the same bins as in
    Figure~\ref{fig:hardall}. Yellow boxes denote SPARQL endpoints,
    green dataset descriptions, blue generic information resources and
    red vocabularies.}
  \label{fig:hardtable}
\end{figure}

We also note that prohibition of SPARQL results caching is rare.
Amongst the servers that expose caching headers it is common
that the result may be cached for some minutes.

\subsubsection{Simple heuristic freshness estimates}\label{sec:simplefresh}

We next consider the simple heuristic freshness lifetime as suggested in
Section~4.2.2 in RFC7234 and mentioned in the introduction.

\begin{figure}[th!]
  \centerline{%
    \includegraphics[width=.9\textwidth]{heuristicall.pdf}}
  \caption{Barplot counting all simple heuristic freshness lifetimes
    found. Bins as in Figure~\ref{fig:hardall}.}
  \label{fig:heuristicall}
\end{figure}

We were able to compute a heuristic lifetime for 554 resources, a
larger number than standards-compliant resources. In
Figure~\ref{fig:heuristicall}, we see that the distribution of
lifetimes are radically different from the case in
Figure~\ref{fig:hardall}. In this case, we may cache many resources
for months. Only a handful of resources had changed in the last
minutes. Since this is based on actual times since last modifications,
this suggests that many resources should have had explicit cache
headers with very long lifetimes. This is supported by
DyLDO\cite{dyldo2}, which concludes that:
\begin{quote}
[\ldots] We found that 62.2\% of documents didn’t change over the six
months and found that 51.9\% of domains were considered static.
\end{quote}
This agrees well with that 60\% of the simple heuristic lifetimes are
in the month range.

Moreover, by inspecting Figure~\ref{fig:heuristictable}, we
note that the difference between different types of resources is much
smaller. This is confirmed by a hypothesis test that yields \pvalue =
0.02.

\begin{figure}[hb!]
  \centerline{%
    \includegraphics[width=.9\textwidth]{heuristictable.pdf}}
  \caption{Mosaic Plot for heuristic freshness lifetime. See caption
    of Figure~\ref{fig:hardtable} for description.  }
  \label{fig:heuristictable}
\end{figure}

We find that only one SPARQL endpoint yields a heuristic lifetime, on
closer inspection, we find this to be hosted by
Dydra\footnote{\url{http://dydra.com/}}. We speculate that this is due to
that few underlying DBMS systems help track modification times in a
way that can be used on a SPARQL result basis.

\subsubsection{Heuristic freshness from Dublin Core properties}

We noted that the Dublin Core Metadata terms vocabulary has a number
of predicates that may become useful in determining heuristic
freshness in the future, so we recorded any statements containing the
predicates \rdfterm{dct:date}, \rdfterm{dct:accrualPeriodicity},
\rdfterm{dct:created}, \rdfterm{dct:issued}, \rdfterm{dct:modified} or
\rdfterm{dct:valid}.

First, we compare dates given in \rdfterm{dct:modified} to dates given
in \httph{Last-Modified} when both are available for a resource. They
are often not the same, but it appears that dates in the former are
further back in time than the latter. We speculate that this may be
due to that the web-server tracks semantically insignificant changes
through the file system, while authors of RDF only update timestamp
when significant changes are made, or it may be that authors forget to
update their timestamps.

\rdfterm{dct:modified} occurs in 2687 triples, but 2487 of these does
not have the Request-URI of the information resource as their subject,
i.e. it gives the last modification time of some subgraph of the
returned RDF. Nevertheless, given its prevalence, it is highly likely
that the presence of \rdfterm{dct:modified} will be useful in
determining a heuristic freshness lifetime, as the latest date may be
used.

\rdfterm{dct:valid} occurred 21 times, and could be used in lieu of an
\httph{Expires} header, but none of these occurrences had a date in the
future.

\rdfterm{dct:accrualPeriodicity} occurred only twice, and in neither
case contained machine readable data.

\rdfterm{dct:date}, \rdfterm{dct:created} and \rdfterm{dct:issued}
were present in 36, 389 and 1475 triples respectively. They correspond
roughly to the \httph{Date} header, which is present in all requests,
and they are therefore not important, but given their prevalence they
could be useful in further analysis.

\subsection{Cache validation}

Hitherto, we have considered the case where the client or proxy does
not send a HTTP request to the server when a resource that is present
in the cache is requested. This is desirable if the client can be
sufficiently confident that the response is fresh, either by having a
standards-compliant freshness lifetime or a heuristic to determine
it. At the end of this period, or if the previous response had no
lifetime, due to lack of information, or to that the server has stated
so in the \httph{Cache-Control} header, responses must be
revalidated. In this case, RFC7232 defines the behaviour, with
\httph{ETag} and \httph{Last-Modified} as the relevant response
headers.

The BTC had recorded these headers in their data, where 1733 had the
\httph{ETag} header and 690 had \httph{Last-Modified} with a great
overlap. For the resources where either or both were available, we
made our initial request conditional, and 911 responses were verified
as still fresh.

In total, 1260 successful initial responses contained
an \httph{ETag}, 606 for vocabularies, 117 for datasets, just 12 for
endpoints and 525 for unclassified information resources.

To see if the server actually supported conditional requests, and not
just merely set the response headers, we made another 1822 requests to
the resources that had these headers. Then, we checked if the response
code was 200, and the conditional headers had not changed since our
initial requests. In 85 cases, conditional requests were not supported
according to the standard, no cases for endpoints, 3 for datasets and
23 for vocabularies, 59 for generic information resources.

\section{Conclusions and outlook}

We found moderate uptake for HTTP caching and conditional requests in
the Semantic Web. We found, in agreement with DyLDO\cite{dyldo2}, that
many resources change at a very slow pace, but also that this is not
reflected in the standards-compliant freshness lifetimes advertised by
servers.

We found that errors are commonplace, but that they do not usually
pertain to the caching headers. We found a small number of
self-contradictory \httph{Cache-Control} headers, and some servers
that set conditional request response headers, but could not support
conditional requests.

It is possible in a substantial number of cases to compute a heuristic
freshness lifetime, either from the \httph{Last-Modified} header, or
from Dublin Core properties.

For SPARQL endpoints, we found that conditional requests are 
seldomly supported, but standards-compliant freshness lifetimes have
been seen, and since it is supported by DBPedia, benefits from caching
may be realised already.

In spite of this, most of the Semantic Web is unhelpful even though it
is changing slowly.

\subsection{Future work}

In this work, we have only explored the cases where the server is
cooperative, in the sense that message data or metadata provides at
least a hint of a resource's cacheability. We also noted that the
majority of the Semantic Web is unhelpful. Therefore, an interesting
direction is to learn the change frequency of resources to use in a
heuristic freshness lifetime estimation. However, such work should
operate within the loose constraints of Section~4.2.2 in RFC7234, and
should find its niche when the other techniques described in this paper
are unavailable. Once this is done, the correctness of caching headers
should be assessed, possibly using contingency tables similar to those
in the companion technical report \cite{kjernsmo_add_survey_2015}.

Investigate whether curated collections such as LOV or SPARQLES
contain resources that have different characteristics. Since the
different sources we surveyed overlap, this cannot be done with the
simple hypothesis test in this paper, but requires more sophisticated
statistics.

We found that some implementations are likely better than others, but
further understanding of these differences are impeded by the fact
that most \httph{Server} headers contained very little
information. Future work could seek more sophisticated fingerprinting
and understanding of the nature of these differences. With this, it
may also be possible to investigate whether expiry times have been set
consciously.

Further investigate the suitability of the \rdfterm{dct:modified}
property for estimating heuristic freshness lifetime.

Estimating the freshness lifetime is a challenging problem for data
owners. It must necessarily involve the human users involved in the
publishing cycle since they are making a commitment about future
changes. Designing user support systems as well as interfaces that fit
the publisher's workflow is an important problem.

For SPARQL endpoints, it is interesting to examine the queries in a
shared proxy, e.g. an institutional cache or a Content
Delivery Network for frequent subpatterns, and prefetch and cache such
intermediate results so that subsequent queries can be answered in
part by the proxy.

We believe that  \cite{ldf1} and \cite{hogan2014paths} prematurely
reject caching in connection to SPARQL. They are correct that
currently, query results can only be cached on a per-query
basis. Moreover, semantically insignificant changes to a query, such
as whitespace or the order of triple patterns in a basic graph
pattern, currently causes problems for caches. The latter problem can
probably be fixed by developing digest algorithms. Such algorithms
exist, but are focused on cryptographical strength, and much simpler
algorithms could be used for this problem.

Furthermore, by using similarity measures, like those discussed in
\cite{dividino2013following}, proxies can proactively prefetch data
into a cache to help answer queries. A cost model that takes into
account several data access methods, those described by
\cite{hogan2014paths} as well as  the novel work in \cite{ldf1} may be
a key ingredient in enabling efficient SPARQL evaluation on the Web.

Even though it is clear that other parts of the Web benefit greatly
from caching, and that the potential for HTTP metadata to better
reflect actual update practices in the Semantic Web is great,
estimating the actual impact of doing so should be done.

\subsection{Recommendations}

It is mainly the objective of this paper to be descriptive, but from
the results of this study and on the results of DyLDO, we note that it
is highly likely that most cache prohibitions are misguided, and
server administrators are advised to turn them off unless they are
sure they are required.

Additionally, based in part on the survey, but also on other practical
experience, we suggest the following:

In many cases, setting reasonable cache headers is straightforward,
and should be done by data owners. Framework authors should make it
easy to set the expected lifetime, heeding the metadata association
good practice recommendation of \cite{Jacobs:04:AWW}. If the author is
unable to influence HTTP headers, they should set a
\rdfterm{dct:valid} time into the future and make use of
\rdfterm{dct:modified}.

To allow generation and validation of \httph{Last-Modified} and
\httph{ETag} headers, DBMS authors should make sure it is much cheaper
to retrieve the modification time of any subgraph, than to retrieve
the subgraph itself. This would be a great improvement for
RFC7232-based caching, when revalidation is required. It would also
help simple heuristics based caching. Research in that direction has
been published in \cite{kaseicache}.

A change periodicity predicate should be standardized in VoID \cite{voidnote}.

All Web cache implementations we have studied have cached responses in
the form of a key that identifies a serialized object. For short-term
impact, future work should accept this as an architectural constraint.

\paragraph*{Acknowledgements} The author would like to thank Jürgen
Umbrich and Martin Giese for careful review and critical comments, and
Gregory Todd Williams for promptly solving issues in the underlying
libraries used in this study, Axel Polleres for encouraging comments
and Jonas Smedegaard for proofreading. We also thank the anonymous
reviewers for extensive reviews.

\bibliographystyle{plain}
%\bibliographystyle{abbrv}
%\bibliographystyle{jbact}
%\bibliographystyle{splncs03}
\bibliography{egne,webarch,data,rfc,hypermedia,dynamicity,stat1,specs,vocabs,similarity,optimization}

\end{document}