### 2015-03-13 sanity/letter.tex
Start of response letter
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@1082 43ef1190-90fd-4dd7-a769-d0121f02b80e

\documentclass{article}
\usepackage[english,british]{babel}

\usepackage[utf8]{inputenc}
\usepackage{cite}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{appendix}

\newcommand{\rdfterm}[1]{\texttt{#1}}
\newcommand{\pvalue}{\textit{p}-value\ }
\newcommand{\httph}[1]{\texttt{#1}}
\newcommand{\todo}[1]{\ensuremath{^{\textrm{\tiny{TODO}\normalsize}}}\footnote{\textbf{TODO:}~#1}}

\title{Letter concerning changes in ``A survey of HTTP caching implementations on the open Semantic Web''}
\author{Kjetil Kjernsmo}

%\subtitle{---Unpublished Working Draft, do not circulate---}

\begin{document}
\maketitle

I have as requested made extensive changes to the submitted
paper. First, the appendices were moved to a Technical Report, due to
be published by the University of Oslo shortly as
\cite{kjernsmo_add_survey_2015}. This report will also include some
extra material, particularly statistical methods, that were removed
from the paper before submission.

\section{Meta-review items}

\begin{enumerate}

\item We include Table~2, which lists the \httph{Server} headers for
  hosts that enabled a freshness lifetime to be computed for
  \emph{all} requests. It makes the conclusion that not much can be
  found from these headers clear, and shows how they should look. For
  the contingency table test, the reader is referred to
  \cite{kjernsmo_add_survey_2015}.

\item Number of occurences is now explained in the caption. The
  percentage numbers have been hacked to display well.

\item This comment seems to pertain to different aspects of the paper:
  First, I acknowledge that I quote many numbers in the text, outside
  of the descriptive statistics, and I understand that this may seem
  disorganized, and that the desire may emerge to see them nicely
  tabulated. They don't lend themselves well to tabular display, since
  they come from quite different parts of the analysis. For example, I
  don't think it would serve paper well if the figures of the
  \httph{Pragma} headers would be in a table with the
  \httph{Cache-Control} headers, the discussion is too
  different. Also, I note that the majority of the reviewers rated the
  paper well with respect to clarity. The other aspect of the paper
  seems related to coverage. To answer that, I have noted that the
  relative sizes of the databases are not important. See also the next
  item.

\item I have made a separate section entitled ``Challenges to
  validity'' to address some of these concerns. In retrospect, using
  LODstats for a list of vocabularies would probably have spared me
  much trouble, but should not challenge the validity much.

\item First, it is clearly out of the scope of the paper to discuss
  how to cache and get a performance improvement. As this is the focus
  of my current work, I took the opportunity to talk about it dearly
  anyway, and extended the future work section with a discussion. I
  note that the reviewer had changed their review to ask us to
  ``acknowledge the complexity to cache SPARQL'', which I think I had
  done in the related work section where I wrote that ``In some of the
  related work, it is shown that caching does not necessarily give
  tangible benefits.'', but I didn't at that time find space to
  elaborate. With space freed by the technical report, I have now done
  so anyway.

\item A very brief comment in ``Future work'', as doing this as a
  major undertaking. Ultimately, it can probably only be done by an
  extensive survey using questionnaires.

\item I acknowledge the error, and reran the spider recording
  \httph{Surrogate-Capability}, which is the correct header, but that
  were not observed either.

\item First, I changed ``RFC7234, Section~4.2.2'' to ``Section~4.2.2
  in RFC7234'', which I hope should clarify that it pertains to the
  standard. Second, since it is my agenda to spread knowledge about
  foundational Web technologys, I would like to follow this
  suggestion. However, to satisfy all the other requirements, I found
  this hard to prioritize, after all this is about the HTTP standard,
  a standard that is fundamental to all practical contributions to the
  Semantic Web. Since review 5 requested further elaboration on how
  the revalidation process works, I detailed that. I also reintroduced
  the table of all recorded headers, that previously was in the
  appendix.

\item See Section~\ref{rev6}.

\item I have made another pass of the paper, and also sent it to a
  professional English proofreader, but don't know if it will fit in
  their schedule to review it before the deadline.

\end{enumerate}

\section{Reviewer 6's comments}\label{rev6}

\begin{description}
\item [Contribution list] I have rewritten two paragraphs, essentially
  keeping the objectives, but detailing how meeting those objectives
  realize the contributions.
\item [Preliminaries and Fundamentals] I have, as noted about, further
  expanded on the RFCs, and put that in a separate subsection under
  Introduction, as I think the text flows better with that
  arrangement. I also included the table as per their suggestion.
\item [Methodology] I identified the the source of the correlation
  that reviewer mentioned as due to a single host having many SPARQL
  endpoints configured similarly and noted that.
\item [Data] I have included a URL to my homepage, where I will make
  an archive available shortly with data and code (I have to admit
  that the code isn't all that pretty, I didn't intend the survey to
  become as extensive as it that, that's why I didn't just release it
  immediately).
\item [Analysis1] I'm not quite sure what they are getting at, but
  included a paragraph just after the section headline.
\item [Analysis2] See Section~\ref{histogramstuff}.
\item [Discussion] This is really about an extensive restructuring of
  the paper, see Section~\ref{restructure} for my take.
\end{description}

%\bibliographystyle{plain}
\bibliographystyle{abbrv}
%\bibliographystyle{jbact}
%\bibliographystyle{splncs03}
\bibliography{egne,webarch,data,rfc,hypermedia,dynamicity,stat1,specs,vocabs,similarity,optimization}

\end{document}