### 2016-06-05 dissertation/conclusions.tex
some conclusions #9

\section{Conclusions}\label{sec:conclusions}

\begin{itemize}
\item Hypermedia is one key to developing applications on the Semantic
  Web.
\item Traits eases the task of developing experimental systems and
  usage of optimisations in underlying systems.
\item While the adoption of caching and conditional request as defined
  in HTTP is not widespread, the adoption is sufficient to start using
  it in practical systems.
\item Evaluation methodologies suffer from a poor foundation in
  philosophy of science, and research into the epistemology of the
  field is important.
\item The author admits that the study is lacking in terms of
  objectivity, but notes that the problems are of a general nature.
\item Statistical Design of Experiments provides a path to a critical
  practice of evaluations, needed to improve the foundations of the
  field.
\end{itemize}