### 2016-06-23 dissertation/introduction.tex
Turtle example

\chapter{Introduction}

The World Wide Web, or just the Web for short, is a well-known global
information space invented by Tim Berners-Lee in 1989, further emerged
in the 1990-ties. It is characterised first and foremost by its
universality, anyone can set up a computer, connect it to the Internet and
start serving data or documents from it, and further adapt it to their purpose.

The Semantic Web is an extension of the Web that has been under
development since 1997\footnote{The first working group draft of the
  RDF specification is dated 1997-08-01, see
  \url{http://www.w3.org/TR/WD-rdf-syntax-971002/}.}, to extend the
Web with languages for expressing information in a machine processable
form\cite{semwebroadmap}.

The Semantic Web is built on a number of specifications that has been
developed under the auspices of the World Wide Web Consortium. The
core technology is known as the Resource Description Framework
(RDF). The framework is defined in a suite of specifications, defining
concepts and abstract syntax, semantics, and several concrete
syntaxes.

Conceptually, an RDF statement is a triple, where the terms are a
subject, a predicate and an object. The building blocks for these
triples are literals, blank nodes and Internationalized Resource
Identifiers (IRI). IRIs builds on the familiar URL: Uniform Resource
Locator, informally known as a Web address,
e.g. \url{https://www.w3.org/RDF/} is an address for a document about
RDF. On the Semantic Web, the URL is first generalised to a Uniform
Resource Identifier, URI, that can not only name a document on the
Web, but any resource. A resource can be a physical thing, like a
person, or an abstract thing, like the type of weather at some place
at some time, numbers, strings, etc. Then, URIs are extended to IRIs
to enable the use of any character in them. Literals are strings with
a datatype (denoted by a IRI), and optionally a language tag. Finally,
blank nodes are unnamed nodes. RDF defines which of these building
blocks may be used for what term: The subject may be a IRI or a blank
node, the predicate only a IRI, whereas the object may be a literal,
blank node or IRI.

The use of IRIs makes it possible to view RDF statements as a directed
graph, where the subject IRI is a vertex, the predicate is an edge to
an object vertex\footnote{In general it is more precisely described as a
  partially labelled directed pseudograph with unique node labels}.

For historical reasons, the first syntax that was standardised for RDF
was an syntax based on the tree-structured Extensible Markup Language
(XML). The simplest standardised syntax is known as N-Triples, where
each statement is written on a single line, where each IRI is written
out in full. The basic building blocks of N-Triples has been reused to
create a compact and natural text form syntax, known as Turtle. Turtle
has gained traction, and is well established as a common syntax. A
simple graph expressed in Turtle could be:

\begin{verbatim}
@prefix foaf: <http://xmlns.com/foaf/0.1/>.
@prefix dct:  <http://purl.org/dc/terms/>.

<http://www.kjetil.kjernsmo.net/foaf#me> a foaf:Person ;
    foaf:name "Kjetil Kjernsmo" ;
    foaf:publications <http://folk.uio.no/kjekje/2013/iswc.pdf> .
<http://folk.uio.no/kjekje/2013/iswc.pdf> a foaf:Document ;
    dct:title "Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation" ; 
    dct:creator <http://www.kjetil.kjernsmo.net/foaf#me>, 
                <http://data.semanticweb.org/person/john-s-tyssedal> .
\end{verbatim}

The example opens with prefix declarations, to allow abbreviating
IRIs, e.g. the predicate \rdfterm{foaf:name} is an abbreviation of the
IRI \rdfterm{http://xmlns.com/foaf/0.1/name}. The IRI
\rdfterm{http://www.kjetil.kjernsmo.net/foaf#me} is a name for this
author. The \rdfterm{a} predicate is standards-defined abbreviation
for \rdfterm{http://www.w3.org/1999/02/22-rdf-syntax-ns#type}, which
is used for declaring the subject an instance of a class, in this case
\rdfterm{foaf:Person}. Then, the \rdfterm{foaf:publications} predicate
is used to link to an author-published version of one of the
publications of this dissertation, and further information is then
provided about this publication, including a link to the co-author. As
we see, semicolons are used to separate statements that share a
subject, commas separate statements with the same subject and
predicate, while periods terminate a statement.

Also, note that all IRIs in this example are \emph{dereferenceable},
meaning that if an HTTP client is used, a representation of the
resource can be retrieved across the Internet. This, along with that
it is expressed as RDF, makes the above so-called Linked Data.

The present work sits in the confluence of several contemporary
efforts in the Semantic Web community. The focus is on query answering
with the SPARQL query language, with emphasis on exploiting the World
Wide Web, but it touches upon query federation, hypermedia, empiricial
methods for evaluating performance, standards compliance and even
philosophy of science.

\section{Contributions}

\subsection{Hypermedia}

\begin{enumerate}
\item A thorough discussion of the implications of Hypermedia
  Types\cite{hypermediatypes} for RDF.
\item A sketch of a vocabulary for read-write RDF hypermedia.
\end{enumerate}

\subsection{Design of Experiments}

\begin{enumerate}
\item Introduction of a path to critical practice of evaluations, that makes
  use of contemporary statistical techniques, to establish a practice
  that can be used to refute assertions on performance.
\item A didactical experiment to help researchers understand the statistics.
\item The novel application of a well established method in
  statistics, rarely used in Computer Science, to SPARQL endpoint
  evaluation.

\end{enumerate}

\subsection{Development problems}

\begin{enumerate}
\item A framework to enable the use of low-level optimisations in
  databases.
\item Simplification when implementing experimental features in
  SPARQL.
\item Make it possible to compose custom query planners that doesn't
  require inheritance.
\end{enumerate}

\subsection{Survey of HTTP Caching}

\begin{enumerate}
\item An understanding of actual usage of caching headers (metadata)
  on the open Semantic Web.
\end{enumerate}

\subsection{Philosophy of Science}

\begin{enumerate}
\item A provocation to discuss epistemological questions.
\end{enumerate}