### 2016-06-24 dissertation/details.tex
HTTP details

\section{Preliminaries}

\subsection{Caching in the Hypertext Transfer Protocol Standard}\label{sec:httpcache}

Internet standards come in the shape of Requests for Comments (RFC),
that it is hoped will gain traction to become \textit{de facto}
standards. Two such RFCs regulate caching in the Internet when HTTP is
used: RFC7234 ``Caching'' \cite{rfc7234} and RFC7232 ``Conditional
Requests'' \cite{rfc7232}. In \cite{kjernsmo_survey_2015}, I provided
some details on these standards in Section~1.1, that is worth
reiterating to understand the problems and contributions of this
work.

RFC7234 defines how clients and proxies may reuse a prior response
without contacting the origin server at all. The server can do this by
supplying a freshness lifetime. It also defines when clients and/or
proxies may \emph{not} reuse a prior response, and what can be done
with stale responses. It also contains a Section~4.2.2 titled
``Calculating Heuristic Freshness'', which sets some loose constraints
for when a prior response may be used even when the server has not
supplied a freshness lifetime. These constraints include that
heuristics may not be used if there is an explicit freshness lifetime,
or the server prohibits caching. The section also proposes a simple
heuristic for calculating freshness lifetime.

RFC7232 defines a protocol for asking the origin server if the cached
response is still fresh. Two response headers are relevant to achieve
this: \httph{ETag} and \httph{Last-Modified}.  \httph{ETag} may
contain an opaque identifier for a specific version of a resource
representation, for example a query result. A client or proxy may then
ask the origin server if the resource is still fresh, by sending a
request with a \httph{If-None-Match} header. If the server is able to
validate the resource representation, it will respond with a 304
status code, otherwise it will return an updated representation. It
may be of use if servers are able to generate and validate such opaque
identifiers cheaply.

The \httph{Last-Modified} header should contain the time of last
modification of the resource representation. A similar protocol is
used to validate it.