### 2016-07-24 dissertation/details.tex
details on cartesians

\section{Preliminaries}

This section extends the basic introduction in
Section~\ref{sec:semwebtech} where required to understand the
following discussion.

\subsection{Caching in the Hypertext Transfer Protocol Standard}\label{sec:httpcache}

Internet standards come in the shape of Requests for Comments (RFC),
that it is hoped will gain traction to become \textit{de facto}
standards. Two such RFCs regulate caching in the Internet when HTTP is
used: RFC7234 ``Caching'' \cite{rfc7234} and RFC7232 ``Conditional
Requests'' \cite{rfc7232}. In \cite{kjernsmo_survey_2015}, I provided
some details on these standards in Section~1.1, that is worth
reiterating to understand the problems and contributions of this
work.

RFC7234 defines how clients and proxies may reuse a prior response
without contacting the origin server at all. The server can do this by
supplying a freshness lifetime. It also defines when clients and/or
proxies may \emph{not} reuse a prior response, and what can be done
with stale responses. It also contains a Section~4.2.2 titled
``Calculating Heuristic Freshness'', which sets some loose constraints
for when a prior response may be used even when the server has not
supplied a freshness lifetime. These constraints include that
heuristics may not be used if there is an explicit freshness lifetime,
or the server prohibits caching. The section also proposes a simple
heuristic for calculating freshness lifetime.

RFC7232 defines a protocol for asking the origin server if the cached
response is still fresh. Two response headers are relevant to achieve
this: \httph{ETag} and \httph{Last-Modified}.  \httph{ETag} may
contain an opaque identifier for a specific version of a resource
representation, for example a query result. A client or proxy may then
ask the origin server if the resource is still fresh, by sending a
request with a \httph{If-None-Match} header. If the server is able to
validate the resource representation, it will respond with a 304
status code, otherwise it will return an updated representation. It
may be of use if servers are able to generate and validate such opaque
identifiers cheaply.

The \httph{Last-Modified} header should contain the time of last
modification of the resource representation. A similar protocol is
used to validate it.

\subsection{Topics in query evaluation}\label{sec:prelimquery}

As shown in \cite{perez2009semantics} that SPARQL has some highly
complex parts. In particular, nested \sparql{OPTIONAL}s will be
difficult to evaluate. It is possible for a user to either inadvertently
or with malicious intent (in e.g. an denial of service attack) submit
a query that will cause excessive load to the server.

Another problem worthy of special attention is the case where parts of
the query has no connecting variable. Let us consider the
simple example of when two triple patterns do not share a variable:
\begin{verbatim}
?person foaf:name ?name .
?contribution dct:title ?title .
\end{verbatim}
When this is evaluated, it will cause what is known as a ``Cartesian
join'', in which the result will contain \emph{all} the solutions of
the first triple pattern joined with \emph{all} the solutions of the
second triple pattern. This will usually be a large result, and it is
also resource consuming to compute.

In some cases, the users may have written such a query, in which case
the query planner only has the options to execute the query or reject
it entirely. In other cases, the query planner may arrive at such
queries as part of query rewriting, in which case it must ensure that
such plans are not preferred.