### 2016-07-30 dissertation/index.tex

\documentclass[a4paper, 12pt]{uiophd}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{multirow}
\usepackage{listings}
\usepackage{supertabular}
\usepackage{cite}
\usepackage{amsthm}
\usepackage{tikz}
\usepackage{pdfpages}
\usepackage[toc]{appendix}
\usepackage{todonotes}
\usepackage{pdflscape}
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}

\include{notasjon}
\include{helpers}
\usepackage[pdftex,linktocpage]{hyperref}
\hyperbaseurl{https://github.com/}

\newtheorem{problem}{Problem}
\newtheorem{example}{Example}

\begin{document}
\title{SPARQL on the Open, Decentralised Web}
\author{Kjetil Kjernsmo}
%\institute{Department of Informatics,
%Postboks 1080 Blindern,
%0316 Oslo, Norway}

%\email{kjekje@ifi.uio.no}

\frontmatter
\maketitle

\input{abstract}

\tableofcontents
\mainmatter

\input{introduction}

\input{introdefhistory}
\input{motivation}

\chapter{Technology Details}
\input{details}
\input{related}

\chapter{Problems and Solutions}

\input{problems}
\input{papersincontext}
\input{conclusions}
\input{cache-proxy}

\bibliography{management,dataprofiles,federation,dynamicity,hypermedia,specs,webarch,practical,semweb,caching,critisism,data,philosophy,benchmarks,rfc,programming,egne,optimization}
\bibliographystyle{plain}

\chapter{Papers}

\includepdf[pages=1-last,openright]{/home/kjetil/uio/www_docs/2012/hypermedia-rdf.pdf}
\includepdf[pages=1-last,openright]{/home/kjetil/uio/www_docs/2013/iswc.pdf}
\includepdf[pages=1-last,openright]{paper14.pdf}
\includepdf[pages=1-last,openright]{/home/kjetil/uio/www_docs/2015/cache-survey.pdf}
\includepdf[pages=1-last,openright]{report.pdf}
\includepdf[pages=1-last,openright]{noise.pdf}
\end{document}