### 2016-07-30 dissertation/acknowledgments.tex
and Helen

\section*{Acknowledgements}

First and foremost it is my pleasure to thank my caring wife Hege,
whose invaluable support and affection has been critical to carry
through this project, and then my children Synne, Eivind and Marius,
two of whom have been born while working on this, for putting up with
a father that sometimes has his mind in the computer. My deceased
parents Ragnhild and Dag has had a major impact with their stimulus
throughout childhood and their enduring support. Arguably, allowing me
to use the 1985 PC with no graphics adapter to write programs was the
start of this project.

My advisers Martin Giese and Arild Waaler deserves a big thanks for
kickstarting the project, help adjusting its course over the years and
thorough and enlightened criticism of manuscripts prior to submission.

I would like to thank my co-authors: In particular Gregory Todd
Williams, who have not only co-authored a paper, but written much of
the underlying software I use, formed the standards that this work
builds on, and been a great discussion partner and friend. John
S. Tyssedal deserves a big thanks for agreeing to write a paper where
we started out with very little common ground. Discussions with Ruben
Verborgh always creates the right amount sparks to be enlightening,
and it has been a pleasure to become friends.

It is also a pleasure to thank the entire Semantic Web community for
their openness and welcoming attitude. Whether you come as a hacker
or want to pursue standardisation, you are likely to meet people like
Dan Brickley, Charles Nevile, Andy Seaborne, Steve Harris, Phil Archer
or Ivan Herman, all of whom have been important for my understanding
of the Semantic Web. The academic community, with their important ISWC
and ESWC conferences are equally welcoming, and has culture of sharing
insight that many other academic communities would envy. Many thanks
are due to Oscar Corcho for first integrating me into the community,
Maria-Esther Vidal and Jacco van Ossenbruggen for encouraging me to
further pursue the speculations around the role of empirical methods,
and general support. Jürgen Umbrich and Axel Polleres deserves a big
thanks for welcoming me in Vienna and many discussions.

I have made many friends in the community, for which I am thankful. Of
these, the following in no particular order have in one way or the
other shaped this work through personal contact beyond published
papers: Olaf Görlitz, Norman Gray, Ansgar Scherp, Frank Van Harmelen,
Olaf Hartig, Javier David Fernández García, Aidan Hogan, Maribel
Acosta and Philip Stutz.

The Perl community have also been important in developing the ideas
and the code of this thesis. In addition to Gregory Todd Williams, I
would like to thank Toby Inkster, Chris Prather, Kip Hampton, Robin
Berjon, Shawn M.  Moore, Michael Nachbaur and Jonas Smedegaard.

My industry colleagues David Norheim and Robert Engels have
also been important in shaping my technical understanding and my
ideas.

Finally, many thanks to Helen Murray for linguistic assistance.