### 2016-11-16 dissertation/index.tex
List of papers in Paper Chapter, fixes #29

\documentclass[a4paper, 12pt, twoside]{uiophd}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{multirow}
\usepackage{listings}
\usepackage{supertabular}
\usepackage{cite}
\usepackage{amsthm}
\usepackage{tikz}
\usepackage{framed}
\usepackage{pdfpages}
\usepackage[toc]{appendix}
\usepackage{todonotes}
\usepackage{pdflscape}
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}

\include{notasjon}
\include{helpers}
\PassOptionsToPackage{hyphens}{url}\usepackage[pdftex,linktocpage]{hyperref}
\hyperbaseurl{https://github.com/}
\hypersetup{  
  pdfinfo={  
    Title={SPARQL on the Open, Decentralised Web},  
    Author={Kjetil Kjernsmo}
  }  
} 
\newtheorem{problem}{Problem}
\newtheorem{example}{Example}

\begin{document}
\title{SPARQL on the Open, Decentralised Web}
\author{Kjetil Kjernsmo}
%\institute{Department of Informatics,
%Postboks 1080 Blindern,
%0316 Oslo, Norway}

%\email{kjekje@ifi.uio.no}

\frontmatter
\maketitle
\newpage
This page was intentionally left blank
\newpage
\input{abstract}
\input{acknowledgments}

\tableofcontents
\mainmatter

\input{introduction}

\input{introdefhistory}
\input{motivation}

\chapter{Preliminaries and Related Work}
\input{details}
\input{related}

\chapter{Problems and Solutions}

\input{problems}
\input{papersincontext}
\input{conclusions}
\input{cache-proxy}

\bibliography{management,dataprofiles,federation,dynamicity,hypermedia,specs,webarch,practical,semweb,caching,critisism,data,philosophy,benchmarks,rfc,programming,egne,optimization}
\bibliographystyle{plain}

\chapter{Papers}

The following papers are used for this thesis:
\begin{itemize}
\item Kjetil Kjernsmo.
\newblock The Necessity of Hypermedia RDF and an Approach to Achieve it.
\newblock In {\em Proceedings of the First Linked APIs workshop at the Ninth
  Extended Semantic Web Conference}, May 2012.
\item Kjetil Kjernsmo and John~S. Tyssedal.
\newblock Introducing Statistical Design of Experiments to SPARQL Endpoint
  Evaluation.
\newblock In Harith Alani, Lalana Kagal, Achille Fokoue, Paul Groth, Chris
  Biemann, Josiane~Xavier Parreira, Lora Aroyo, Natasha Noy, Chris Welty, and
  Krzysztof Janowicz, editors, {\em The Semantic Web – ISWC 2013}, volume
  8219 of {\em Lecture Notes in Computer Science}, pages 360--375. Springer
  Berlin Heidelberg, 2013.
\item Gregory~Todd Williams and Kjetil Kjernsmo.
\newblock Pushing Complexity Down the Stack.
\newblock In {\em Proceedings of the ISWC Developers Workshop 2014}, volume
  1268 of {\em {CEUR} Workshop Proceedings}. CEUR-WS.org, May 2014.
\item Kjetil Kjernsmo.
\newblock A Survey of HTTP Caching Implementations on the Open Semantic Web.
\newblock In Fabien Gandon, Marta Sabou, Harald Sack, Claudia d’Amato,
  Philippe Cudré-Mauroux, and Antoine Zimmermann, editors, {\em The Semantic
  Web. Latest Advances and New Domains}, volume 9088 of {\em Lecture Notes in
  Computer Science}, pages 286--301. Springer International
  Publishing, 2015.
\item Kjetil Kjernsmo.
\newblock Addendum to a survey of {HTTP} caching on the {Semantic Web}.
\newblock Technical Report 444, Department of Informatics, University of Oslo,
  March 2015.
\item Kjetil Kjernsmo and Ruben Verborgh.
\newblock How can scientific methods provide guidance for semantic web research
  and development.
\newblock In {\em Workshop on Negative or Inconclusive Results in Semantic Web
  Proceedings}, volume 1435 of {\em {CEUR} Workshop Proceedings}. CEUR-WS.org,
  2015.
\end{itemize}

\includepdf[pages=1-last,openright]{/home/kjetil/uio/www_docs/2012/hypermedia-rdf.pdf}
\includepdf[pages=1-last,openright]{/home/kjetil/uio/www_docs/2013/iswc.pdf}
\includepdf[pages=1-last,openright]{paper14.pdf}
\includepdf[pages=1-last,openright]{/home/kjetil/uio/www_docs/2015/cache-survey.pdf}
\includepdf[pages=1-last,openright]{report.pdf}
\includepdf[pages=1-last,openright]{noise.pdf}
\end{document}